---
title: "Principle component analysis vs support vector machine"
author: "Dong Liang"
date: '2018-12-02'
output:
  html_document:
    df_print: paged
  pdf_document: 
    fig_caption: yes
    keep_tex: yes
  word_document: default
editor_options:
  chunk_output_type: inline
slug: priciple-component-analysis-vs-support-vector-machine
tags: Machine learning
categories: Machine learning
---

## Background

This spreadsheet  contains urine proteomic data normalized to creatinine (relative fluorescent units of protein per mg of creatinine) with 4 groups: 1) Febrile UTI), 2) nonfebrile UTI, 3) culture negative pyuria (white blood cells in the urine) and 4) culture negative normal UA

We are looking for a panel of  protein/creatinine levels that predicts UTI (group 1 and 2 combined) when compared to both culture negative pyuria  (group 3) and culture negative, normal UA (group 4) individually and in combination


```{r}
library(knitr)
opts_chunk$set(engine.path = '/anaconda3/bin/python')
```

## Load proteomics data

```{r, engine='python'}
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
# plt.style.use('ggplot')
import matplotlib
plt.style.use('ggplot')
# matplotlib.use('Agg')
scan = pd.read_csv('/Users/DL/Documents/Jupyter project/Somascan Protein Creatinine/somascan.csv')
print(scan.head())
```

## Exploratory data analysis

## Data processing 
Covert pandas dataframe to matrix.
We build an internal `Data2matrix` class for `Pipeline` to use. We will need implement fit and transform functions inside the Data2matrix class.

```{r, engine='python'}
from sklearn.preprocessing import Imputer, StandardScaler, MinMaxScaler
from sklearn.pipeline import Pipeline
from sklearn.base import BaseEstimator, TransformerMixin
from sklearn.model_selection import cross_val_score, cross_val_predict
# imputer = Imputer(strategy = 'mean')

class DataFrameSelector(BaseEstimator, TransformerMixin):
    def __init__(self, attribute_names):
        self.attribute_names = attribute_names #[attribute_names != 'sampleID']
    def fit(self, X, y = None):
        return self
    def transform(self, X):
        return X[self.attribute_names].values
    

class Data2matrix(BaseEstimator, TransformerMixin):
    def __init__(self, attribute_names = None):
        self.attribute_names = attribute_names 
    def fit(self, X, y = None):
        return self
    def transform(self, X):
        return X.values
        

num_pipeline = Pipeline([
    ('Data2matrix', Data2matrix()),
    ('std_scalar', StandardScaler()) ])

```

## Prepare feature X matrix and label y
```{r, engine='python'}
proteinCreatine = scan.T
X = proteinCreatine.iloc[3::, :]
Xz = (X - X.mean(axis = 0))/X.std(axis=0)
Xz = Xz.values

# Prepare feature X matrix
X = num_pipeline.fit_transform(X)

# Prepare label y
g1 = ['febrile'] * 8
g2 = ['afebriel'] * 8
g3 = ['negPyuria'] * 8
g4 = ['negNoPyuria'] * 8
# y = [*g1, *g2, *g3, *g4]
```

```{r, engine='python'}
u, s, vh = np.linalg.svd(X)
pc = X.dot(vh)
print(pc[:, 0:3])

plt.scatter(pc[:, 0], pc[:, 1])
plt.show()
```


